---
title: "Communities"
format: html
---

# Open Source Communities

Browse through our comprehensive list of open source communities. Use the filters below to find communities by region or category.

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(DT)
library(htmltools)

# Read communities data
communities <- read_csv("../data/communities.csv")

# Make website URLs clickable with proper format
communities <- communities %>%
  mutate(website_display = ifelse(
    startsWith(website, "http"),
    paste0("<a href='", website, "' target='_blank'>", website, "</a>"),
    paste0("<a href='https://", website, "' target='_blank'>", website, "</a>")
  ))

# Select only the columns we want to display
display_communities <- communities %>%
  select(name, description, website_display, region, category, format)

# Create format options for filtering
format_options <- communities %>%
  distinct(format) %>%
  pull()
```

## Community Directory

```{r}
#| label: datatable
#| output: html
datatable(
  display_communities,
  filter = list(
    position = "top",
    clear = FALSE,
    plain = FALSE
  ),
  options = list(
    pageLength = 10,
    autoWidth = FALSE,
    scrollX = FALSE,
    columnDefs = list(
      list(width = '20%', targets = 0),  # name
      list(width = '30%', targets = 1),  # description
      list(width = '15%', targets = 2, 
           render = JS("function(data, type, row) { return type === 'display' ? data : data.replace(/<.*?>/g, ''); }")),  # website
      list(width = '12%', targets = 3),  # region
      list(width = '13%', targets = 4),  # category
      list(width = '10%', targets = 5)   # format
    ),
    # Custom filter for the format column
    initComplete = JS("
      function(settings, json) {
        var table = this.api();
        
        // Create select filter for format column (6th column, index 5)
        var formatColumn = table.column(5);
        var formatSelect = $('<select><option value=\"\">All</option><option value=\"in-person\">In-Person</option><option value=\"virtual\">Virtual</option><option value=\"hybrid\">Hybrid</option></select>')
          .appendTo($(formatColumn.header()).closest('tr').siblings('.filters').children().eq(5).empty())
          .on('change', function() {
            formatColumn
              .search($(this).val() ? '^' + $(this).val() + '$' : '', true, false)
              .draw();
          });
      }
    ")
  ),
  escape = FALSE  # Important: Allow HTML in the table
)
```

## How to Add Your Community

If your community is not listed, you can add it by:
1. Creating a GitHub Issue using our template
2. Submitting a Pull Request with your community details

See our [Contribution Guidelines](../contribute.html) for more information.